\huge \bf Checkers with AI

### 2014-03-19 paper/paper.tex
Nathan wrote his section.

@@

    \subsection{Haskell}
        \subsubsection{Introduction}
            Unlike C/C++, Python doesn't have built in compatibility with Haskell. Haskell however
            does have a foreign function interface built in to the language and the Glasgow Haskell
            Compiler. There is a number of libraries out there which deal with the Haskell FFI and
            allow easy interfacing with other languages. One such library is called HaPy, which
            provides an interface between Python and Haskell which allows Python to directly call
            Haskell functions, much like how ctypes works with C/C++.
        \subsubsection{Code Example}
            \lstinputlisting[language=Python, firstline=3, lastline=3,
                title=Importing from Python (checkers.py)]{../checkers.py}
            \lstinputlisting[language=Haskell,
                title=Exporting the AI module to HaPy (Export.hs)]{../Export.hs}
            \lstinputlisting[language=Haskell, firstline=1, lastline=7,
                title=First few lines of defining our AI module (HsklAI.hs)]{../HsklAI.hs}
        \subsubsection{How the code works}
            The code and installation instructions for HaPy can be
            found here: \\ \url{https://github.com/sakana/HaPy} \\
            Generally what this library does is that once you compile Haskell into a
            shared library (much like we do for C/C++) the Python types are converted into
            a C representation, and then into a Haskell representation, then the function is
            executed. The return type is propagated in reverse to Python 
            $(Haskell \rightarrow C \rightarrow Python)$.
            HaPy currently supports converting Python primitive types (int, string, etc.) 
            as well as lists into their equivalent in Haskell. Calling the function is also
            abstracted away nicely so that when the programmer wants to call a Haskell 
            function from Python, is looks exactly like a standard python library
            function call would.
            \lstinputlisting[language=Python, firstline=171, lastline=171]{../checkers.py}
        \subsubsection{Why we chose this}
            Using a library like HaPy to do the dirty work of using the Foreign Function Interface
            allowed us to focus more on writing good Haskell code. Instead of worrying about 
            the representation of our types and how they would propagate to Python, we just wrote
            our Haskell code and let the library take care of the type issues. Over all, HaPy is
            easy to set up (at least on the Mac system we primarily tested on) and has a clean
            and easy to use interface on the Python side. It requires a non trivial amount
            of set up to initially get everything to work, but HaPy has clear instructions
            on how to get project files set up. And once the libraries are installed in both
            Python and Haskell everything works beautifully.

>>>>

\section{Algorithms}

	\subsection{Move Validation}
	Move validation was the bulk of our ugly code as the best way of implementing was to hardcode all the conditions that had to be met and matching them. We separated the move validation and application to simply the latter as applying moves was complicated enough in haskell.

	\subsection{Flipping the board}
Early on we decided to simplify the AI's code by always giving the AI the game state as if they were the player sitting at the top of the board. This was accomplished by flipping before the AI was called on the bottom player.		The flipping code had to be implemented in the AI's as well because the MinMax algorithm requires playing as the other player every other depth

	\subsection{MinMax}
The MinMax algorithm consists of finding every possible board position after a set number of moves. Then Generating a score for the boards final position. The board was stored in a tree with each node storing a move, a score, and having a list of children nodes. The tree is generated by recursively calling the function for each possible move with a depth of one less the previous and a version of the board with the move applied and then this board is flipped to simulate the other players turn. The building part recurses until it reaches the maximum depth specified and calculates and saves the heuristic in the leaf node. The function return on leaf node and all the nodes are added to the list in a depth first fashion. After the tree is built it it traversed depth first again but this time will traverse all the way to the bottom non-leaf nodes and find the maximum or minimum heuristic for the node depending on whether it simulated your turn or the opponent. The calls keep returning the max or min until the best value is which ever node has the highest value at the end.

	\subsection{Heuristic}
Our heuristic was chosen due to its simplicity and ease in debugging its decisions. We simply added the pieces on the board with the opponents pieces being negative to the score and your pieces being positive. When summing the regular pieces were worth 1 and kings are worth 2. We also added code to detect a stalemate and made it a maximum value when the opponent cant move and a minimum amount when you cant move since this is a loss for the player that can't win.

	\subsection{Language Specific Deviations}
C++ was the first language we decided to implement and it stuck closest to the design. When we made modifications we changed the design. The calling of C++ code in python with C types made it so we had to insert the move output stored in the STL list type of C++ into a C array so python could read it.
\par
Haskells lack of for loops made the code look different but overall we simply used a map our function over the sequence [0:31] to simulate indexes into arrays. The part of the algorithm that preformed MinMax traversal of the tree was much more complicated due to Haskells lack of states so the algorithm to finds the extreme of the row required a self recursive call as well as a downward depth recursive call and a comparison of all the return values of the depth call. This compares to the C++ code where we only had to recourse downwards and compare the results to a variable stored outside the for loop.
\par
Python only needed the Validate move to checks the moves inputted. The python code for validation move was basically the same as C as it turn out the best method was just a bunch of if statements checking conditions for each movement. To check for stale make all possible a list of all feasible 1 space moves or single jumps was generated and if any passed validating there was no stalemate.

@@

    \subsection{Haskell}
        \subsubsection{Introduction}
            Unlike C/C++, Python doesn't have built in compatibility with Haskell. Haskell however
            does have a foreign function interface built in to the language and the Glasgow Haskell
            Compiler. There is a number of libraries out there which deal with the Haskell FFI and
            allow easy interfacing with other languages. One such library is called HaPy, which
            provides an interface between Python and Haskell which allows Python to directly call
            Haskell functions, much like how ctypes works with C/C++.
        \subsubsection{Code Example}
            \lstinputlisting[language=Python, firstline=3, lastline=3,
                title=Importing from Python (checkers.py)]{../checkers.py}
            \lstinputlisting[language=Haskell,
                title=Exporting the AI module to HaPy (Export.hs)]{../Export.hs}
            \lstinputlisting[language=Haskell, firstline=1, lastline=7,
                title=First few lines of defining our AI module (HsklAI.hs)]{../HsklAI.hs}
        \subsubsection{How the code works}
            The code and installation instructions for HaPy can be
            found here: \\ \url{https://github.com/sakana/HaPy} \\
            Generally what this library does is that once you compile Haskell into a
            shared library (much like we do for C/C++) the Python types are converted into
            a C representation, and then into a Haskell representation, then the function is
            executed. The return type is propagated in reverse to Python 
            $(Haskell \rightarrow C \rightarrow Python)$.
            HaPy currently supports converting Python primitive types (int, string, etc.) 
            as well as lists into their equivalent in Haskell. Calling the function is also
            abstracted away nicely so that when the programmer wants to call a Haskell 
            function from Python, is looks exactly like a standard python library
            function call would.
            \lstinputlisting[language=Python, firstline=171, lastline=171]{../checkers.py}
        \subsubsection{Why we chose this}
            Using a library like HaPy to do the dirty work of using the Foreign Function Interface
            allowed us to focus more on writing good Haskell code. Instead of worrying about 
            the representation of our types and how they would propagate to Python, we just wrote
            our Haskell code and let the library take care of the type issues. Over all, HaPy is
            easy to set up (at least on the Mac system we primarily tested on) and has a clean
            and easy to use interface on the Python side. It requires a non trivial amount
            of set up to initially get everything to work, but HaPy has clear instructions
            on how to get project files set up. And once the libraries are installed in both
            Python and Haskell everything works beautifully. 
            %We could have used Sockets or stdin instead of the hooking the guctions into python but each require a lot more bookkeeping code and add there own complications.
